Constructive Preference Elicitation by Setwise Max-margin Learning

### 2016-01-25 paper/ijcai16.tex
paper: fixed clash between feature and query number

\paragraph{Setting.} We assume to have a multi-attribute feature space $\calX$
of configurations $\vx = (x_1, \ldots, x_n)$ over $n$ features. For the time
being we will focus on binary features only, i.e. $x_z\in\{0,1\}$ for all
$z\in[n]$, to keep matters simple. This is a common choice for preference
elicitation methods~\cite{guo2010real,viappiani2010optimal}; discrete features
can be handled for instance by means of a simple one-hot encoding. Support for
linearly dependent continuous features will be discussed later on. 
\andrea{mmh sembra che gestiamo solo booleani.. riscriverei che ci concentriamo solo su feature categoriche, e 
che le rappresentiamo con one-hot encoding come tipico dei metodi preference elicitation..}

@@

In adherence to previous work~\cite{guo2010real,viappiani2010optimal}, we model users by means
of linear utility functions: the user's preferences are represented by a weight
vector $\vw\in\bbR^n$, and the utility of a configuration $\vx$ is given by
$\langle \vw, \vx \rangle = \sum_{z=1}^n w_z x_z$. In the remainder of the paper we
require all weights to be {\em non-negative} and {\em bounded}: the per-attribute
weights $w_z$ must lie in a (constant but otherwise arbitrary) interval
$[w^\bot_z, w^\top_z]$, with $w^\bot_z \ge 0$. Both requirements are quite
natural, and enable the translation of our core optimization problem into a
mixed-integer linear problem (as done in Section~\ref{sec:formulation}).

@@

A straightforward encoding of the above desiderata leads to the
following mixed integer {\em non-linear} optimization problem over the
variables $M$, $\{ \vw^i \}$, $\{ \vx^i \}$: \andrea{dovremmo forse specificare i domini delle variabili per chiarire la natura mixed}
%
{\footnotesize
\begin{align}
    \max
        & \;\; M - \alpha \sum_{i=1}^k \| \veps^{i} \|_1 - \beta \sum_{i=1}^k \| \vw^{i} \|_1 + \gamma \sum_{i=1}^k \langle \vw^{i}, \vx^{i} \rangle
        \nonumber
    \\
    \text{s.t.}
        & \;\; \forall \; i \in [k], \forall \; h \in [n] \nonumber
    \\
        & \;\; \qquad \langle \vw^{i}, \vy^{h}_+ - \vy^{h}_- \rangle \ge M - \varepsilon^{i}_h \label{eq:wyconstr}
    \\
        & \;\; \forall \; i, j \in [k], i \neq j \;.\; \langle \vw^{i}, \vx^{i} - \vx^{j} \rangle \ge M \label{eq:wxconstr}
    \\
        & \;\; \forall \; i \in [k] \;.\; \vw^\bot \le \vw^{i} \le \vw^\top \label{eq:wbounds}
    \\
        & \;\; \forall \; i \in [k] \;.\; \vx^{i} \in \calX_{\text{feasible}} \label{eq:xbounds}
    \\
        & \;\; \forall \; i \in [k] \;.\; \veps^{i} \ge 0 \nonumber
    \\
        & \;\; M \ge 0 \nonumber
\end{align}
} \andrea{perche' un punto tra il quantificatore e la formula? inoltre mi sa che ``n'' e' usato per due cose, queries e features}
%
Let us illustrate the above piece by piece. The objective is composed of four
parts: we maximize the shared margin $M$ (first part) and minimize the total
sum of the ranking errors $\veps^i$ incurred by each weight vector $\vw^{i}$
(second part), while at the same time regularizing the magnitude of the
weights (third part) and the quality of the configurations $\{ \vx^{i} \}$ (last
part). The non-negative hyperparameters $\alpha,\beta,\gamma$ control the
influence of the various components.

@@

In order to do so, we introduce a set of fresh variables $p^{i,j}_z$ for every
$i,j\in[k]$ and $z\in[n]$. Assuming for the time being that the new variables
do satisfy the equation $p^{i,j}_z = w^i_z x^j_z$, we rewrite the fourth
component of the objective function in terms of the new variables as:
%
$$ \gamma \sum_{i=1}^k \sum_{z=1}^n p^{i,i}_z $$
%
and, similarly, Eq.~(\ref{eq:wxconstr}) as:
%
\[ \forall \; i, j \in [k], i \neq j \;.\; \sum_{z=1}^n p^{i,i}_z - p^{i,j}_z \ge M \label{eq:pxconstr} \]
%
The fact that $p^{i,j}_z = w^{i}_z x^{j}_z$ is achieved by
setting the following additional constraints. We distinguish between two cases:
(i) $p^{i,i}_z$ and (ii) $p^{i,j}_z$ for $i \ne j$.  Recall that we are
maximizing the margin $M$. Now, due to Eq.~(\ref{eq:pxconstr}), the optimizer will
try to keep $p^{i,i}_z$ as large as possible and $p^{i,j}_z$ as small as
possible.

@@

Substituting the above MILP constraints into the original non-linear
formulation, we obtain the following mixed-integer linear problem:
%
{\footnotesize
\begin{align}
    \max
        & \;\; M - \alpha \sum_{i=1}^k \| \veps^{i} \|_1 - \beta \sum_{i=1}^k \| \vw^{i} \|_1 + \gamma \sum_{i=1}^k \sum_{z=1}^n p^{i,i}_z
        \nonumber
    \\
    \text{s.t.}
        & \;\; \forall \; i \in [k], \forall \; h \in [n] \nonumber
    \\
        & \;\; \qquad \langle \vw^{i}, \vy^{h}_+ - \vy^{h}_- \rangle \ge M - \varepsilon^{i}_h \nonumber
    \\
        & \;\; \forall \; i, j \in [k], i \neq j \;.\; \sum_{z=1}^n p^{i,i}_z - p^{i,j}_z \ge M
    \\
    %     & \;\; \forall \; i \in [k], \forall \; z \in [n] \nonumber
    % \\
        & \;\; \forall \; i, j \in [k], i \neq j, \forall \; z \in [n] \nonumber
    \\
        & \;\; \qquad p^{i,i}_z \le \min \{ w_\text{max} x^{i}_z, w^{i}_z \}
    \\
        & \;\; \qquad p^{i,j}_z \ge \max \{ 0, w^{i}_z - w_\text{max}(1 - x^{j}_z) \}
    \\
        & \;\; \forall \; i \in [k] \;.\; \vw^\bot \le \vw^{i} \le \vw^\top \label{eq:wbounds2}
    \\
        & \;\; \forall \; i \in [k] \;.\; \vx^{i} \in \calX_{\text{feasible}} \nonumber
    \\
        & \;\; \forall \; i \in [k] \;.\; \veps^{i}_h \ge 0 \nonumber
    \\
        & \;\; M \ge 0 \nonumber
\end{align}
}
%
which can be solved by any suitable MILP solver.

@@

\paragraph{Synthetic Dataset.} Following the experimental protocol
in~\cite{guo2010real} and \cite{viappiani2010optimal}, in the first experiment
we evaluate the behavior of the proposed method in a challenging artificial
setting. We developed seven synthetic datasets: each dataset involves
$n=2,\ldots,7$ attributes, where each attribute takes one of $n$ possible
values. More explicitly, for $n=2$ the synthetic dataset is
$\calX_\text{feasible} = [2] \times [2]$, for $n=3$ to $\calX_\text{feasible} =
[3] \times [3] \times [3]$, and so on. The cardinality of
$\calX_\text{feasible}$ is $n^n$, and grows (super) exponentially with $n$. For
$n=3$, the dataset is comparable to the synthetic one used
in~\cite{guo2010real} and~\cite{viappiani2010optimal} (which have three domains
of size of 2, 2 and 5, respectively, for a total of 20 feasible
configurations), while for larger $n$ it grows much larger than those typically
used in the Bayesian preference elicitation literature, and as such represents
a good test bed for comparing the scalability of the various methods.
The feasible configuration space was encoded in {\sc SetMargin} through
appropriate MILP constraints, while the other methods required all datasets to
be explicitly grounded.

>>>>

\paragraph{Setting.} We assume to have a multi-attribute feature space
$\calX$ of configurations $\vx = (x_1, \ldots, x_m)$ over $m$
features. For the sake of simplicity we focus on binary features only,
i.e. $x_z\in\{0,1\}$ for all $z\in[m]$, assuming a one-hot encoding of
categorical features. This is a common choice for preference
elicitation methods~\cite{guo2010real,viappiani2010optimal}. Support
for linearly dependent continuous features will be discussed later on.

@@

In adherence to previous work~\cite{guo2010real,viappiani2010optimal}, we model users by means
of linear utility functions: the user's preferences are represented by a weight
vector $\vw\in\bbR^m$, and the utility of a configuration $\vx$ is given by
$\langle \vw, \vx \rangle = \sum_{z=1}^m w_z x_z$. In the remainder of the paper we
require all weights to be {\em non-negative} and {\em bounded}: the per-attribute
weights $w_z$ must lie in a (constant but otherwise arbitrary) interval
$[w^\bot_z, w^\top_z]$, with $w^\bot_z \ge 0$. Both requirements are quite
natural, and enable the translation of our core optimization problem into a
mixed-integer linear problem (as done in Section~\ref{sec:formulation}).

@@

A straightforward encoding of the above desiderata leads to the
following mixed integer {\em non-linear} optimization problem over the
variables $M$, $\{ \vw^i \}$, $\{ \vx^i \}$: \andrea{dovremmo forse specificare i domini delle variabili per chiarire la natura mixed}
%
{\footnotesize
\begin{align}
    \max
        & \;\; M - \alpha \sum_{i=1}^k \| \veps^{i} \|_1 - \beta \sum_{i=1}^k \| \vw^{i} \|_1 + \gamma \sum_{i=1}^k \langle \vw^{i}, \vx^{i} \rangle
        \nonumber
    \\
    \text{s.t.}
        & \;\; \forall \; i \in [k], \forall \; h \in [n] \nonumber
    \\
        & \;\; \qquad \langle \vw^{i}, \vy^{h}_+ - \vy^{h}_- \rangle \ge M - \varepsilon^{i}_h \label{eq:wyconstr}
    \\
        & \;\; \forall \; i, j \in [k], i \neq j \;.\; \langle \vw^{i}, \vx^{i} - \vx^{j} \rangle \ge M \label{eq:wxconstr}
    \\
        & \;\; \forall \; i \in [k] \;.\; \vw^\bot \le \vw^{i} \le \vw^\top \label{eq:wbounds}
    \\
        & \;\; \forall \; i \in [k] \;.\; \vx^{i} \in \calX_{\text{feasible}} \label{eq:xbounds}
    \\
        & \;\; \forall \; i \in [k] \;.\; \veps^{i} \ge 0 \nonumber
    \\
        & \;\; M \ge 0 \nonumber
\end{align}
} \andrea{perche' un punto tra il quantificatore e la formula?}
%
Let us illustrate the above piece by piece. The objective is composed of four
parts: we maximize the shared margin $M$ (first part) and minimize the total
sum of the ranking errors $\veps^i$ incurred by each weight vector $\vw^{i}$
(second part), while at the same time regularizing the magnitude of the
weights (third part) and the quality of the configurations $\{ \vx^{i} \}$ (last
part). The non-negative hyperparameters $\alpha,\beta,\gamma$ control the
influence of the various components.

@@

In order to do so, we introduce a set of fresh variables $p^{i,j}_z$ for every
$i,j\in[k]$ and $z\in[m]$. Assuming for the time being that the new variables
do satisfy the equation $p^{i,j}_z = w^i_z x^j_z$, we rewrite the fourth
component of the objective function in terms of the new variables as:
%
$$ \gamma \sum_{i=1}^k \sum_{z=1}^m p^{i,i}_z $$
%
and, similarly, Eq.~(\ref{eq:wxconstr}) as:
%
\[ \forall \; i, j \in [k], i \neq j \;.\; \sum_{z=1}^m p^{i,i}_z - p^{i,j}_z \ge M \label{eq:pxconstr} \]
%
The fact that $p^{i,j}_z = w^{i}_z x^{j}_z$ is achieved by
setting the following additional constraints. We distinguish between two cases:
(i) $p^{i,i}_z$ and (ii) $p^{i,j}_z$ for $i \ne j$.  Recall that we are
maximizing the margin $M$. Now, due to Eq.~(\ref{eq:pxconstr}), the optimizer will
try to keep $p^{i,i}_z$ as large as possible and $p^{i,j}_z$ as small as
possible.

@@

Substituting the above MILP constraints into the original non-linear
formulation, we obtain the following mixed-integer linear problem:
%
{\footnotesize
\begin{align}
    \max
        & \;\; M - \alpha \sum_{i=1}^k \| \veps^{i} \|_1 - \beta \sum_{i=1}^k \| \vw^{i} \|_1 + \gamma \sum_{i=1}^k \sum_{z=1}^m p^{i,i}_z
        \nonumber
    \\
    \text{s.t.}
        & \;\; \forall \; i \in [k], \forall \; h \in [n] \nonumber
    \\
        & \;\; \qquad \langle \vw^{i}, \vy^{h}_+ - \vy^{h}_- \rangle \ge M - \varepsilon^{i}_h \nonumber
    \\
        & \;\; \forall \; i, j \in [k], i \neq j \;.\; \sum_{z=1}^m p^{i,i}_z - p^{i,j}_z \ge M
    \\
    %     & \;\; \forall \; i \in [k], \forall \; z \in [m] \nonumber
    % \\
        & \;\; \forall \; i, j \in [k], i \neq j, \forall \; z \in [m] \nonumber
    \\
        & \;\; \qquad p^{i,i}_z \le \min \{ w_\text{max} x^{i}_z, w^{i}_z \}
    \\
        & \;\; \qquad p^{i,j}_z \ge \max \{ 0, w^{i}_z - w_\text{max}(1 - x^{j}_z) \}
    \\
        & \;\; \forall \; i \in [k] \;.\; \vw^\bot \le \vw^{i} \le \vw^\top \label{eq:wbounds2}
    \\
        & \;\; \forall \; i \in [k] \;.\; \vx^{i} \in \calX_{\text{feasible}} \nonumber
    \\
        & \;\; \forall \; i \in [k] \;.\; \veps^{i}_h \ge 0 \nonumber
    \\
        & \;\; M \ge 0 \nonumber
\end{align}
}
%
which can be solved by any suitable MILP solver.

@@

\paragraph{Synthetic Dataset.} Following the experimental protocol
in~\cite{guo2010real} and \cite{viappiani2010optimal}, in the first experiment
we evaluate the behavior of the proposed method in a challenging artificial
setting. We developed seven synthetic datasets: each dataset involves
$m=2,\ldots,7$ attributes, where each attribute takes one of $m$ possible
values. More explicitly, for $m=2$ the synthetic dataset is
$\calX_\text{feasible} = [2] \times [2]$, for $m=3$ to $\calX_\text{feasible} =
[3] \times [3] \times [3]$, and so on. The cardinality of
$\calX_\text{feasible}$ is $m^m$, and grows (super) exponentially with $m$. For
$m=3$, the dataset is comparable to the synthetic one used
in~\cite{guo2010real} and~\cite{viappiani2010optimal} (which have three domains
of size of 2, 2 and 5, respectively, for a total of 20 feasible
configurations), while for larger $m$ it grows much larger than those typically
used in the Bayesian preference elicitation literature, and as such represents
a good test bed for comparing the scalability of the various methods.
The feasible configuration space was encoded in {\sc SetMargin} through
appropriate MILP constraints, while the other methods required all datasets to
be explicitly grounded.
